---
title: "MovieMining"
author: "Rita Zou, Fei Xu, Yang Li, Xinyu Cao, Rus Refati"
date: "11/28/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Packages (not sure what is needed)

```{r}
library(tidyverse)
library(dplyr)
library(caret)
library(lubridate)
```

Read in Data, remove text heavy columns

### Preprocessing

```{r}
#setwd("~/Desktop/UTK/BZAN 542 - Data Mining/Group Project")
MOVIES <- read.csv("imdb_top_1000.csv")

 ########### Filter Data ########### 
#Remove poster link and plot overview
MOVIES <- MOVIES %>% select(-c(Poster_Link,Overview))
#Store all blanks as NA
MOVIES[MOVIES==""] <- NA
#Remove rows with NAs for Gross:
MOVIES <- MOVIES[-which(is.na(MOVIES$Gross)),]

 ########### Fix Data Types ########### 
#Released_year
  MOVIES[804,"Certificate"] <- MOVIES[804,"Released_Year"]; MOVIES[804,"Released_Year"] <- "1995" #Year courtesy of Google
  MOVIES$Released_Year <- as.numeric(MOVIES$Released_Year)
#Runtime
  MOVIES$Runtime <- as.integer(sapply(strsplit(MOVIES$Runtime," "), '[',1))
#Certificate
  MOVIES$Certificate <- as.factor(MOVIES$Certificate)
#Certificate
  MOVIES$Genre <- as.factor(MOVIES$Genre)
#Gross
  MOVIES$Gross <- as.numeric(gsub(",","",MOVIES$Gross)) 
#Stars
  for(i in grep("Star", names(MOVIES))){
    MOVIES[,i] <- as.factor(MOVIES[,i])
  }
 ########### Create Indicator Variables ########### 
  
    ### Certificate ###
    MOVIES <- MOVIES[-which(is.na(MOVIES$Certificate)),]
    formula <- as.formula( paste("~",names(MOVIES)[3],"-1",sep="") ) #set up formula for model.matrix 
    INDICATORS <- data.frame(model.matrix(formula,data=MOVIES))
    MOVIES <- cbind(MOVIES,INDICATORS)
    MOVIES <- MOVIES[,-3]
    
   ### Genre ###
# First step: Get all of the genres that occur in the data
AllGenres <- unique(levels(as.factor(MOVIES$Genre)))
ListOfGenres <- c()
for(i in AllGenres){
  templist <- unlist( strsplit(i, " " ) )
  for(j in 1:length(templist)){
   tempgenre <- unlist(strsplit(templist[j], ",") )
    if( tempgenre %in% ListOfGenres ) { next }
   else {ListOfGenres[length(ListOfGenres)+1] <- tempgenre}
  }
}
length(ListOfGenres)
#21 different genres found

# Second step: Create data frame of indicator variables using found genres
GENREINDICATOR <- as.data.frame(matrix(rep(0, nrow(MOVIES)*length(ListOfGenres)), ncol=length(ListOfGenres)))
names(GENREINDICATOR) <- ListOfGenres #Initialize empty dataframe of all 0's
#Loop through each movie and change the column genre for each one that movie belongs
for(i in 1:nrow(MOVIES)){
  targetmovie <- unlist(strsplit(as.character(MOVIES[i,"Genre"]), " "))
  for(j in 1:length(targetmovie)){
    targetgenre <- unlist(strsplit(targetmovie[j], ","))
    GENREINDICATOR[i,targetgenre] <- 1
  }
}
MOVIES <- cbind(MOVIES, GENREINDICATOR)
 ############################################    



#From Dr. Petrie:
#Looking for zero variance or near zero variance predictors
infodensity <- nearZeroVar(MOVIES, saveMetrics= TRUE)
infodensity[infodensity$nzv,][1:5,]  #dataframe of measures 
#freqRatio = # times most common element appears / # times second most common element appears
#percentUnique = 0-100 percent of distinct values in data; 100 = all values different
zero.variance <- which( nearZeroVar(MOVIES, saveMetrics= TRUE)$zeroVar == TRUE )
near.zero.variance <- which( nearZeroVar(MOVIES, saveMetrics= TRUE)$nzv == TRUE )
#0 variables with zero variance; 16 variables with near-zero variance

#Explore highly correlated or redundant predictors
#Put a minus sign before the column number that contains the y variable
highlycorrelated <- findCorrelation( cor(MOVIES[,which(sapply(MOVIES, class) == "numeric")]) , cutoff = .90)
highlycorrelated # No columns are highly correlated with each other

```

Check for duplicates before renaming rows as Movie Title
```
#### EDIT: After filtering the data with the code above, there are no longer any duplicates ####

# #there is one duplicate in dataset
# length(unique(MOVIES$Series_Title)) 
# nrow(MOVIES)
# 
# MOVIES[which(duplicated(MOVIES$Series_Title)==TRUE),]
# which(MOVIES$Series_Title == "Drishyam")
# MOVIES[c(88,137),] 
# #The two movies were created two years apart in two different languages  -Marc
# 
# MOVIES[88,1] <- "Drishyam1"
# MOVIES[137,1] <- "Drishyam2"
# 
# rownames(MOVIES) <- MOVIES$Series_Title
# MOVIES <- MOVIES %>% select(-(Series_Title))

```

Be it as Gross is so highly skewed, I believe we should take the log

```{r}
hist(MOVIES$Gross)
MOVIES$Gross <- log(MOVIES$Gross)
hist(MOVIES$Gross)
```

Final Check on NA values

```{r}
#check for any NA columns
apply(MOVIES,2,function(x) any(is.na(x)))


#find that metascore does
MOVIES <- MOVIES[-which(is.na(MOVIES$Meta_score) == TRUE),]

summary(MOVIES)
nrow(MOVIES)
```

## Actor / Director Top 50 Rankings

<https://www.ranker.com/crowdranked-list/the-most-oscar-worthy-directors-of-all-time>

<https://www.ranker.com/crowdranked-list/the-greatest-film-actors-and-actresses-of-all-time?ref=collections_page>

Created csv for the top 50 of each, maybe we can decide how to code these into predictors later, either binary or by grouping 0-10,11-20 etc.

Rankings are voted by general population, but are recent and reflect public's view -- which are both positives.

```{r}
RANKINGS <- read.csv("best_actors_directors.csv")
head(RANKINGS)

#check if top50director present
top50director <- c()
for (i in 1:nrow(MOVIES)) {
  if(MOVIES$Director[i] %in% RANKINGS$director){top50director[i] <- 1}
  else{top50director[i] <- 0}
}

MOVIES$top50director <- top50director

#check across all 4 stars if top50 actor present
counter <- c(rep(0,nrow(MOVIES)))
top50actor <- c()
for (i in 1:nrow(MOVIES)) {
  if(MOVIES$Star1[i] %in% RANKINGS$actor){counter[i] <- counter[i]+1}
  else{counter[i] <- counter[i]+0}
  if(MOVIES$Star2[i] %in% RANKINGS$actor){counter[i] <- counter[i]+1}
  else{counter[i] <- counter[i]+0}
  if(MOVIES$Star3[i] %in% RANKINGS$actor){counter[i] <- counter[i]+1}
  else{counter[i] <- counter[i]+0}
  if(MOVIES$Star4[i] %in% RANKINGS$actor){counter[i] <- counter[i]+1}
  else{counter[i] <- counter[i]+0}
  if(counter[i] >= 1){top50actor[i] <- 1}
  else{top50actor[i] <- 0}
}
MOVIES$top50actor <- top50actor


#fix row indices
rownames(MOVIES) <- c(seq(from = 1,to = 714,by=1))

```

## Ultimate Task: Predicting Gross Box Office Sales

Clustering IMDB v Gross

```{r}

set.seed(542)
grpRatings <- kmeans(MOVIES[,c("IMDB_Rating","Gross")],centers = 3, nstart = 50)
grpRatings

min(MOVIES$IMDB_Rating)

plot(MOVIES$IMDB_Rating, MOVIES$Gross, type="n", xlim=c(7,10), xlab="IMDB Rating", ylab="Gross log(Sales)")
text(x=MOVIES$IMDB_Rating, y=MOVIES$Gross, labels=rownames(MOVIES), col=grpRatings$cluster)

MOVIES[1,]

```

Clustering Meta_Score v Gross

```{r}

set.seed(542)
grpRatingsMeta <- kmeans(MOVIES[,c("Meta_score","Gross")],centers = 3, nstart = 50)
grpRatingsMeta

min(MOVIES$Meta_score)

plot(MOVIES$Meta_score, MOVIES$Gross, type="n", xlab="Meta_score", ylab="Gross log(Sales)")
text(x=MOVIES$Meta_score, y=MOVIES$Gross, labels=rownames(MOVIES), col=grpRatingsMeta$cluster)

MOVIES[40,]
```

Clustering Runtime v Gross

```{r}
which(is.na(MOVIES$Runtime) == TRUE)
which(is.na(MOVIES$Gross) == TRUE)

set.seed(542)
grpRunGross <- kmeans(MOVIES[,c("Runtime","Gross")],centers = 3, nstart = 50)
grpRunGross

plot(MOVIES$Runtime, MOVIES$Gross, type="n", xlab="Runtime", ylab="Gross log(Sales")
text(x=MOVIES$Runtime, y=MOVIES$Gross, labels=rownames(MOVIES), col=grpRunGross$cluster)
```

Clustering Certificate v Gross

```{r}

which(is.na(MOVIES$Certificate) == TRUE)
which(is.na(MOVIES$Gross) == TRUE)

CERTIFICATECLUSTER <- MOVIES[-c(which(is.na(MOVIES$Gross) | is.na(MOVIES$Certificate))),]

CERTIFICATECLUSTER$Gross <- as.numeric(gsub(",","",CERTIFICATECLUSTER$Gross)) #Convert gross to numeric

###Because "Certificate" is a categorical variable, we cannot use it in a clustering algorithm directly. We also cannot replace them with values 1,2,3,4, etc.
#Below is the code that will convert each unique Certificate into its own binary indicator variable (courtesy of Dr. Petrie)
CERTIFICATECLUSTER$Certificate <- as.factor(CERTIFICATECLUSTER$Certificate) #Must be a factor
formula <- as.formula( paste("~",names(CERTIFICATECLUSTER)[i],"-1",sep="") ) #set up formula for model.matrix 

INDICATORS <- data.frame(model.matrix(formula,data=CERTIFICATECLUSTER))
CERTIFICATECLUSTER <- cbind(CERTIFICATECLUSTER,INDICATORS) #add columns for the indicator variables

levels(CERTIFICATECLUSTER$Certificate) #There are a lot of different Certificates, and some of them probably mean the same thing. We may want to consider combining some 

set.seed(542)
grpRunGross <- kmeans(MOVIES[,c(13:26)],centers = 3, nstart = 50)
grpRunGross

#Not sure how to plot this...

```

## Discretize Runtime + Gross

```{r}

library(Hmisc)

hist(MOVIES$Gross)

#less than 5 million is a flop
log(5e6)
#5 million to 25million underperforming
log(25e6)
#25million to 100 million mod success
log(100e6)
#100 million to 300 million successful
log(300e6)
#more than 300million is blockbuster



MOVIES <- MOVIES %>% mutate(Box_Office_Groups = case_when(
            Gross <= 15.42495 ~ 1,
            Gross > 15.42495 & Gross <= 17.03439 ~ 2,
            Gross > 17.03439 & Gross <= 18.42068 ~ 3,
            Gross > 18.42068 & Gross <= 19.51929 ~ 4,
            Gross > 19.51929 ~ 5,
))

table(MOVIES$Box_Office_Groups)

MOVIES <- MOVIES %>% mutate(Box_Office_Performance = case_when(
            Box_Office_Groups == 1 ~ "flop" ,
            Box_Office_Groups == 2 ~ "underperforming" ,
            Box_Office_Groups == 3 ~ "moderate success" ,
            Box_Office_Groups == 4 ~ "successful" ,
             Box_Office_Groups == 5 ~ "blockbuster" ,
    )
)

MOVIES <- MOVIES %>% mutate(Runtime_Groups = case_when(
            Runtime <= 90 ~ 1,
            Runtime >90 & Runtime <= 120 ~ 2,
            Runtime >120 & Runtime <= 150 ~ 3,
            Runtime >150 ~ 4,
        )
)
MOVIES <- MOVIES %>% mutate(Runtime_Lengths = case_when(
            Runtime_Groups == 1 ~ "sub1.5" ,
            Runtime_Groups == 2 ~ "1.5to2",
            Runtime_Groups == 3 ~ "2to2.5",
            Runtime_Groups == 4 ~ "2.5andabove",
           
    )
)

#Are these groups significantly different?
#Connecting letters report:
library(multcompView)
AOV <- aov(Gross ~ Box_Office_Performance, data=MOVIES) #Set up left/right side of ~ and data=
TUKEY <- TukeyHSD(AOV)
multcompLetters4(AOV,TUKEY)
#Yes! Each group is significantly different from each other

```

## Test / Train split

```{r}
#80/20 split for industry standard

#use dplyr way
set.seed(542)
index <- sample(1:nrow(MOVIES),size=trunc(.8*nrow(MOVIES)))

MOVIESTRAIN <- MOVIES %>% filter(row_number() %in% index)
MOVIESTEST <- MOVIES %>% filter(!(row_number() %in% index))

length(unique(rownames(MOVIESTRAIN)))
nrow(MOVIESTRAIN)
#wanted to ensure dplyr was sampling w/o replacement...it was

```


## SVM Model

```{r}

# #install.packages('e1071')
# num = sapply(MOVIESTRAIN, is.numeric)  # num contains numeric column indices
# fact = sapply(MOVIESTRAIN, is.factor)
# MOVIESTRAIN_numeric = MOVIESTRAIN[, num]      # the numeric data
# MOVIESTRAIN_factor = MOVIESTRAIN[, fact] 
# drops <- c("CertificateUnrated") #shows CertificateUnrated is constant
# MOVIESTRAIN_numeric_new<-MOVIESTRAIN_numeric[, !(names(MOVIESTRAIN_numeric) %in% drops)]
##
# num1 = sapply(MOVIESTEST, is.numeric)  # num contains numeric column indices
# fact1 = sapply(MOVIESTEST, is.factor)
# MOVIESTEST_numeric = MOVIESTEST[, num]      # the numeric data
# MOVIESTEST_factor = MOVIESTEST[, fact]
# drops <- c("CertificateUnrated")
# MOVIESTEST_numeric_new <-
#   MOVIESTEST_numeric[,!(names(MOVIESTEST_numeric) %in% drops)]
# library(e1071)


svmFit<-train(Box_Office_Performance~.,data=MOVIESTRAIN[,c(2,3,14:49,51)],method="svmLinear",trControl=trainControl(method="cv",number=10))
svmFit
pred<-predict(svmFit,newdata=MOVIESTEST)
table(MOVIESTEST$Box_Office_Performance,pred)# confusion matrix
sum(MOVIESTEST$Box_Office_Performance==pred)/nrow(MOVIESTEST)

svmFit2<-train(Box_Office_Performance~.,data=MOVIESTRAIN[,c(2,3,14:49,51)],method="svmRadial",trControl=trainControl(method="cv",number=10))
svmFit2
pred2<-predict(svmFit2,newdata=MOVIESTEST)
table(MOVIESTEST$Box_Office_Performance,pred2)# confusion matrix
sum(MOVIESTEST$Box_Office_Performance==pred2)/nrow(MOVIESTEST)
```

## Boosted Tree

```{r}
xgbFit<-train(Box_Office_Performance~.,data=MOVIESTRAIN[,c(2,3,14:49,51)],method="xgbTree",
              trControl=trainControl(method="cv",number=10),
              )
xgbFit
```

##Artificial Neural Network

```{r}
nnetFit<-train(Box_Office_Performance~.,data=MOVIESTRAIN[,c(2,3,14:49,51)],method="nnet",trControl=trainControl(method="cv",number=10))
nnetFit
```
